---
title: "RCP 2.6 data processing"
author: "Jiacheng Min"
date: "`r Sys.Date()`"
output: pdf_document
---

## Packages
```{r}
library(readxl)
library(writexl)
library(dplyr)
library(zoo)
library(ggplot2)
library(splines)
library(ISOweek)
library(tidyr)
```

## Load data
```{r}
rcp26.mean = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/UTCI_mean_rcp26.xlsx")[,-1]
rcp26.min = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/UTCI_min_rcp26.xlsx")[,-1]
rcp26.max = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/UTCI_max_rcp26.xlsx")[,-1]

rcp85.mean = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/UTCI_mean_rcp85.xlsx")[,-1]
rcp85.min = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/UTCI_min_rcp85.xlsx")[,-1]
rcp85.max = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/UTCI_max_rcp85.xlsx")[,-1]

```

## Replace NA by Linear Interpolation
```{r}
rcp26.min[-1] <- lapply(rcp26.min[-1], function(col) {
  na.approx(col, rule = 2)  # Linear interpolation
})

rcp26.mean[-1] <- lapply(rcp26.mean[-1], function(col) {
  na.approx(col, rule = 2)  # Linear interpolation
})

rcp26.max[-1] <- lapply(rcp26.max[-1], function(col) {
  na.approx(col, rule = 2)  # Linear interpolation
})

rcp85.min[-1] <- lapply(rcp85.min[-1], function(col) {
  na.approx(col, rule = 2)  # Linear interpolation
})

rcp85.mean[-1] <- lapply(rcp85.mean[-1], function(col) {
  na.approx(col, rule = 2)  # Linear interpolation
})

rcp85.max[-1] <- lapply(rcp85.max[-1], function(col) {
  na.approx(col, rule = 2)  # Linear interpolation
})


# RCP2.6
rcp26.mean$Week <- ISOweek(rcp26.mean$Time)
rcp26.min$Week <- ISOweek(rcp26.min$Time)
rcp26.max$Week <- ISOweek(rcp26.max$Time)


# RCP8.5
rcp85.mean$Week <- ISOweek(rcp85.mean$Time)
rcp85.min$Week <- ISOweek(rcp85.min$Time)
rcp85.max$Week <- ISOweek(rcp85.max$Time)
```


## Coverting the level

The categories relate to UTCI values as follows: 
above +46: extreme heat stress; 4
+38 to +46: very strong heat stress; 3
+32 to +38: strong heat stress; 2
+26 to +32: moderate heat stress; 1
+9 to +26: no thermal stress;  0
+9 to 0: slight cold stress; -1
0 to -13: moderate cold stress; -2
-13 to -27: strong cold stress; -3
-27 to -40: very strong cold stress;  -4
below -40: extreme cold stress. -5

```{r}
classify_utci <- function(x) {
  case_when(
    x < -13  ~ -1,
    x <=  32 ~  0,
    x >   32 ~  1,
    TRUE     ~ NA_real_
  )
}

# 3. Helper to compute weekly heat-/cold-wave counts
summarise_wave <- function(df_max, df_min) {
  hot <- df_max %>%
    mutate(
      Attiki_lvl = classify_utci(Attiki),
      Lisbon_lvl = classify_utci(Lisbon),
      Roma_lvl   = classify_utci(Roma),
      Attiki_hw = as.integer(Attiki_lvl == 1 & lag(Attiki_lvl)==1 & lag(Attiki_lvl,2)==1),
      Lisbon_hw = as.integer(Lisbon_lvl == 1 & lag(Lisbon_lvl)==1 & lag(Lisbon_lvl,2)==1),
      Roma_hw   = as.integer(Roma_lvl   == 1 & lag(Roma_lvl)  ==1 & lag(Roma_lvl,2)==1)
    ) %>%
    replace(is.na(.), 0) %>%
    mutate(Week = ISOweek(Date)) %>%
    group_by(Week) %>%
    summarise(
      Attiki_hot_wave3 = sum(Attiki_hw),
      Lisbon_hot_wave3 = sum(Lisbon_hw),
      Roma_hot_wave3   = sum(Roma_hw),
      .groups = "drop"
    )

  cold <- df_min %>%
    mutate(
      Attiki_lvl = classify_utci(Attiki),
      Lisbon_lvl = classify_utci(Lisbon),
      Roma_lvl   = classify_utci(Roma),
      Attiki_cw = as.integer(Attiki_lvl == -1 & lag(Attiki_lvl)==-1 & lag(Attiki_lvl,2)==-1),
      Lisbon_cw = as.integer(Lisbon_lvl == -1 & lag(Lisbon_lvl)==-1 & lag(Lisbon_lvl,2)==-1),
      Roma_cw   = as.integer(Roma_lvl   == -1 & lag(Roma_lvl)  ==-1 & lag(Roma_lvl,2)==-1)
    ) %>%
    replace(is.na(.), 0) %>%
    mutate(Week = ISOweek(Date)) %>%
    group_by(Week) %>%
    summarise(
      Attiki_cold_wave3 = sum(Attiki_cw),
      Lisbon_cold_wave3 = sum(Lisbon_cw),
      Roma_cold_wave3   = sum(Roma_cw),
      .groups = "drop"
    )

  left_join(hot, cold, by = "Week")
}

```

## Considering the consecutive days
```{r}
# 2. Simplified UTCI classifier
classify_utci <- function(x) {
  case_when(
    x < -13  ~ -1,
    x <=  32 ~  0,
    x >   32 ~  1,
    TRUE     ~ NA_real_
  )
}

# 3. Helper to compute weekly heat-/cold-wave counts
summarise_wave <- function(df_max, df_min) {
  hot <- df_max %>%
    mutate(
      Attiki_lvl = classify_utci(Attiki),
      Lisbon_lvl = classify_utci(Lisbon),
      Roma_lvl   = classify_utci(Roma),
      Attiki_hw = as.integer(Attiki_lvl == 1 & lag(Attiki_lvl)==1 & lag(Attiki_lvl,2)==1),
      Lisbon_hw = as.integer(Lisbon_lvl == 1 & lag(Lisbon_lvl)==1 & lag(Lisbon_lvl,2)==1),
      Roma_hw   = as.integer(Roma_lvl   == 1 & lag(Roma_lvl)  ==1 & lag(Roma_lvl,2)==1)
    ) %>%
    replace(is.na(.), 0) %>%
    mutate(Week = ISOweek(Time)) %>%
    group_by(Week) %>%
    summarise(
      Attiki_hot_wave3 = sum(Attiki_hw),
      Lisbon_hot_wave3 = sum(Lisbon_hw),
      Roma_hot_wave3   = sum(Roma_hw),
      .groups = "drop"
    )

  cold <- df_min %>%
    mutate(
      Attiki_lvl = classify_utci(Attiki),
      Lisbon_lvl = classify_utci(Lisbon),
      Roma_lvl   = classify_utci(Roma),
      Attiki_cw = as.integer(Attiki_lvl == -1 & lag(Attiki_lvl)==-1 & lag(Attiki_lvl,2)==-1),
      Lisbon_cw = as.integer(Lisbon_lvl == -1 & lag(Lisbon_lvl)==-1 & lag(Lisbon_lvl,2)==-1),
      Roma_cw   = as.integer(Roma_lvl   == -1 & lag(Roma_lvl)  ==-1 & lag(Roma_lvl,2)==-1)
    ) %>%
    replace(is.na(.), 0) %>%
    mutate(Week = ISOweek(Time)) %>%
    group_by(Week) %>%
    summarise(
      Attiki_cold_wave3 = sum(Attiki_cw),
      Lisbon_cold_wave3 = sum(Lisbon_cw),
      Roma_cold_wave3   = sum(Roma_cw),
      .groups = "drop"
    )

  left_join(hot, cold, by = "Week")
}

# 4. Compute for each scenario
wave_26 <- summarise_wave(rcp26.max, rcp26.min)
wave_85 <- summarise_wave(rcp85.max, rcp85.min)

# 5. Save into a list and write out
wave_list_rcp26 <- list(
  Attiki = wave_26 %>%
             select(Attiki_hot_wave3,
                    Attiki_cold_wave3),
  Lisbon = wave_26 %>%
             select(Lisbon_hot_wave3,
                    Lisbon_cold_wave3),
  Rome   = wave_26 %>%
             select(Roma_hot_wave3,
                    Roma_cold_wave3)
)

wave_list_rcp85 <- list(
  Attiki = wave_85 %>%
             select(Attiki_hot_wave3,
                    Attiki_cold_wave3),
  Lisbon = wave_85 %>%
             select(Lisbon_hot_wave3,
                    Lisbon_cold_wave3),
  Rome   = wave_85 %>%
             select(Roma_hot_wave3,
                    Roma_cold_wave3)
)

save(
  wave_list_rcp26,
  file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/wave_list_rcp26.RData"
)

save(
  wave_list_rcp85,
  file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/wave_list_rcp85.RData"
)
```

## lag 21 function
```{r}
lag21_function <- function(data_df, region_col, cb_matrix, start_day, fill_na = F) {
  # data_df:      daily data frame containing 'Week' and the region column
  # region_col:   string name of the column to process (e.g., "Attiki")
  # cb_matrix:    pre-allocated matrix (rows = number of weeks, cols = 22)
  # start_day:    start day index of the week (1, 4, or 7)
  
  # define extraction order based on start_day
  start <- switch(as.character(start_day),
    "1" = c(1,7,6,5,4,3,2),
    "4" = c(4,3,2,1,7,6,5),
    "7" = c(7,6,5,4,3,2,1),
    stop("start_day must be 1, 4, or 7")
  )
  
  # group daily values by week into a list-column
  weekly_vals <- data_df %>%
    group_by(Week) %>%
    summarize(vals = list(.data[[region_col]]), .groups = "drop")
  n_weeks <- nrow(weekly_vals)
  if (n_weeks != nrow(cb_matrix)) {
    stop("cb_matrix row count must match number of weeks")
  }
  
  # populate the matrix: current week and lags of 1-2 weeks
  for (i in seq_len(7)) {
    day_vals <- vapply(weekly_vals$vals, `[`, numeric(1), start[i])
    cb_matrix[, i]      <- day_vals
    cb_matrix[, i + 7]  <- dplyr::lag(day_vals, 1)
    cb_matrix[, i + 14] <- dplyr::lag(day_vals, 2)
  }
  # add 3-week lag for the first day only
  cb_matrix[, 22] <- dplyr::lag(vapply(weekly_vals$vals, `[`, numeric(1), start[1]), 3)
  
  # optionally carry nearest non-NA values forward/backward
  if (fill_na) {
    cb_matrix <- apply(cb_matrix, 2, function(col) {
      col <- na.locf(col,      na.rm = FALSE)
      col <- na.locf(col, fromLast = TRUE, na.rm = FALSE)
      col
    })
  }
  
  cb_matrix
}


```

## Crossbasis matrix (not smoothing)
```{r}
n_weeks <- nrow(wave_list_rcp26$Attiki)
Attiki_prep_cb_rcp26 <- matrix(NA_real_, nrow = n_weeks, ncol = 22)

Attiki_prep_cb_rcp26 <- lag21_function(
  data_df    = rcp26.mean,
  region_col = "Attiki",
  cb_matrix  = Attiki_prep_cb_rcp26,
  start_day  = 7,
  fill_na    = TRUE
)

Lisbon_prep_cb_rcp26 <- matrix(NA_real_, nrow = n_weeks, ncol = 22)

Lisbon_prep_cb_rcp26 <- lag21_function(
  data_df    = rcp26.mean,
  region_col = "Lisbon",
  cb_matrix  = Lisbon_prep_cb_rcp26,
  start_day  = 7,
  fill_na    = TRUE
)

Roma_prep_cb_rcp26 <- matrix(NA_real_, nrow = n_weeks, ncol = 22)

Roma_prep_cb_rcp26 <- lag21_function(
  data_df    = rcp26.mean,
  region_col = "Roma",
  cb_matrix  = Roma_prep_cb_rcp26,
  start_day  = 7,
  fill_na    = TRUE
)

n_weeks <- nrow(wave_list_rcp85$Attiki)
Attiki_prep_cb_rcp85 <- matrix(NA_real_, nrow = n_weeks, ncol = 22)

Attiki_prep_cb_rcp85 <- lag21_function(
  data_df    = rcp85.mean,
  region_col = "Attiki",
  cb_matrix  = Attiki_prep_cb_rcp85,
  start_day  = 7,
  fill_na    = TRUE
)

Lisbon_prep_cb_rcp85 <- matrix(NA_real_, nrow = n_weeks, ncol = 22)

Lisbon_prep_cb_rcp85 <- lag21_function(
  data_df    = rcp85.mean,
  region_col = "Lisbon",
  cb_matrix  = Lisbon_prep_cb_rcp85,
  start_day  = 7,
  fill_na    = TRUE
)

Roma_prep_cb_rcp85 <- matrix(NA_real_, nrow = n_weeks, ncol = 22)

Roma_prep_cb_rcp85 <- lag21_function(
  data_df    = rcp85.mean,
  region_col = "Roma",
  cb_matrix  = Roma_prep_cb_rcp85,
  start_day  = 7,
  fill_na    = TRUE
)
```

## Save Crossbasis matrix by RData
```{r}
save(Attiki_prep_cb_rcp26, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/Attiki_cb_rcp26.RData")
save(Lisbon_prep_cb_rcp26, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/Lisbon_cb_rcp26.RData")
save(Roma_prep_cb_rcp26, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp26/Roma_cb_rcp26.RData")

save(Attiki_prep_cb_rcp85, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/Attiki_cb_rcp85.RData")
save(Lisbon_prep_cb_rcp85, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/Lisbon_cb_rcp85.RData")
save(Roma_prep_cb_rcp85, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/Roma_cb_rcp85.RData")
```


## 2031 and 2045 RCP data
```{r}
years    <- c("2031","2045")
measures <- c("mean","min","max")
scenarios<- c("26","85")


for(sc in scenarios){
  for(me in measures){
    # df
    df_name <- paste0("rcp", sc, ".", me)
    df      <- get(df_name)
    for(yr in years){
      # new df
      new_name <- paste0(df_name, ".", yr)
      assign(
        new_name,
        df %>% filter(format(Time, "%Y") == yr)
      )
    }
  }
}
```

## Box-plot
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

make_box_df <- function(measure) {
  bind_rows(
    get(paste0("rcp26.", measure, ".2031")) %>%
      pivot_longer(c(Attiki, Lisbon, Roma),
                   names_to = "Region", values_to = "UTCI") %>%
      mutate(Scenario = "RCP2.6", Year = "2031"),
    get(paste0("rcp85.", measure, ".2031")) %>%
      pivot_longer(c(Attiki, Lisbon, Roma),
                   names_to = "Region", values_to = "UTCI") %>%
      mutate(Scenario = "RCP8.5", Year = "2031"),
    get(paste0("rcp26.", measure, ".2045")) %>%
      pivot_longer(c(Attiki, Lisbon, Roma),
                   names_to = "Region", values_to = "UTCI") %>%
      mutate(Scenario = "RCP2.6", Year = "2045"),
    get(paste0("rcp85.", measure, ".2045")) %>%
      pivot_longer(c(Attiki, Lisbon, Roma),
                   names_to = "Region", values_to = "UTCI") %>%
      mutate(Scenario = "RCP8.5", Year = "2045")
  ) %>%
  mutate(
    Year     = factor(Year,     levels = c("2031","2045")),
    Scenario = factor(Scenario, levels = c("RCP2.6","RCP8.5"))
  )
}

all_min_df  <- make_box_df("min")
all_mean_df <- make_box_df("mean")
all_max_df  <- make_box_df("max")
all_vals    <- c(all_min_df$UTCI, all_mean_df$UTCI, all_max_df$UTCI)
ymin <- floor(min(all_vals, na.rm=TRUE))
ymax <- ceiling(max(all_vals, na.rm=TRUE))

plot_box <- function(measure) {
  df <- make_box_df(measure)
  ggplot(df, aes(x = Year, y = UTCI, fill = Scenario)) +
    geom_boxplot(position = position_dodge(width = 0.8), width = 0.7) +
    facet_wrap(~ Region, nrow = 1, scales = "fixed") +    
    scale_fill_manual(values = c("RCP2.6" = "#4995C6", "RCP8.5" = "#FDA7C9")) +
    scale_y_continuous(limits = c(ymin, ymax)) +          
    labs(
      title = paste0("Daily ", toupper(measure), " UTCI (2031 vs 2045)"),
      x     = "Year",
      y     = paste0(measure, " UTCI"),
      fill  = NULL
    ) +
    theme_minimal(base_size = 14) +
    theme(
      panel.border     = element_rect(color = "black", fill = NA),
      panel.grid.major = element_line(color = "grey90"),
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "grey95", color = NA),
      strip.text       = element_text(face = "bold")
    )
}

p_min  <- plot_box("min")
p_mean <- plot_box("mean")
p_max  <- plot_box("max")

(p_min / p_mean / p_max) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")
```