---
title: "Backfitting DLNM-Lee-Carter Model: Rome"
author: "Jiacheng Min"
date: "`r Sys.Date()`"
output: pdf_document
---


## Used Packages
```{r, warning=FALSE}
packages <- c("readxl", "ggplot2", "forecast", "dplyr", "gridExtra", "writexl", 
              "tseries", "hwwntest", "dlnm", "splines", "ISOweek", "mgcv", 
              "Metrics", "demography", "seastests", "MTS", "tidyr", "grid",
              "nortest")
invisible(lapply(packages, library, character.only = TRUE))
```

## Load data
```{r, warning=FALSE}
setwd("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/")
Y20_64 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y20_64_combined.xlsx")
Y65_74 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y65_74_combined.xlsx")
Y75_84 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y75_84_combined.xlsx")
Y_GE85 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y_GE85_combined.xlsx")
```

## LC matrix (mortality rate), convert to weekly mortality rate
```{r}
Roma = cbind(Y20_64[,6]/Y20_64[,12],Y65_74[,6]/Y65_74[,12], 
             Y75_84[,6]/Y75_84[,12], Y_GE85[,6]/Y_GE85[,12])
colnames(Roma) = c('Y20_64', 'Y65_74', 'Y75_84', 'Y_GE85')
rownames(Roma) = Y20_64$Week
Roma = Roma * 52
```

## Store heat/cold wave data and load crossbasis matrix data
```{r}
Roma_wave = data.frame(cbind(Y20_64$Roma_hot_wave3, Y20_64$Roma_cold_wave3))
colnames(Roma_wave) = c("Roma_hot_wave3", "Roma_cold_wave3")
# crossbasis matrix (before smoothing)
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Roma.RData")



# Smoothing cross-basis matrix (train)
varknots <- equalknots(UTCI_ext.Roma, fun = "ns", df = 3, degree = 3)
lagknots <- logknots(21, 3)
Roma_cb <- crossbasis(UTCI_ext.Roma, lag = 21, 
                      argvar = list(fun = "ns", knots = varknots), 
                      arglag = list(knots = lagknots, df = 3))
  
```

# Load functions
```{r}
files <- list.files("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Function/", pattern = "\\.R$", full.names = TRUE)
sapply(files, source)
```

# Backtesting with expanding window on B-DLNM-LC
```{r, warning=TRUE}
sets = 10
size = 8
forecast_step = 78
start_ind = 102

expanding_window_sets <- lapply(0:(sets-1), function(i) {
  # Training set
  Roma_train <- Roma[1:(start_ind + i * size), ]
  Roma_wave_train <- Roma_wave[1:(start_ind + i * size), ]
  UTCI_ext.Roma_train <- UTCI_ext.Roma[1:(start_ind + i * size), ]
  
  # Smoothing cross-basis matrix (train)
  varknots <- equalknots(UTCI_ext.Roma_train, fun = "ns", df = 3, degree = 3)
  lagknots <- logknots(21, 3)
  Roma_cb_train <- crossbasis(UTCI_ext.Roma_train, lag = 21, 
                                argvar = list(fun = "ns", knots = varknots), 
                                arglag = list(knots = lagknots, df = 3))
  
  # Test set
  Roma_test <- Roma[(start_ind + i * size + 1):(start_ind + i * size + forecast_step), ]
  Roma_wave_test <- Roma_wave[(start_ind + i * size + 1):(start_ind + i * size + forecast_step), ]
  UTCI_ext.Roma_test <- UTCI_ext.Roma[(start_ind + i * size + 1):(start_ind + i * size + forecast_step), ]
  
  # Smoothing cross-basis matrix (test) *** using training set setting ***
  varknots <- equalknots(UTCI_ext.Roma_train, fun = "ns", df = 3, degree = 3)
  lagknots <- logknots(21, 3)
  Roma_cb_test <- crossbasis(UTCI_ext.Roma_test, lag = 21, 
                               argvar = list(fun = "ns", knots = varknots), 
                               arglag = list(knots = lagknots, df = 3)) 
  
  # Return a list of training and test sets for this iteration
  list(
    train = list(Roma_train = Roma_train, Roma_wave_train = Roma_wave_train,
                 Roma_cb_train = Roma_cb_train),
    test = list(Roma_test = Roma_test, Roma_wave_test = Roma_wave_test,
                Roma_cb_test = Roma_cb_test)
  )
})
names(expanding_window_sets) = paste0("set_", 1:sets)

# Apply the B_DLNM_LC fitting and forecasting across all sets
expanding_window_result <- lapply(expanding_window_sets, function(set) {
  # Train the model on the training set
  train_model <- B_DLNM_LC(set$train$Roma_train, 
                           set$train$Roma_cb_train, 
                           set$train$Roma_wave_train, 
                           tol = 1e-2, max_iter = 15, region = "Roma") ## change 
  
  # Forecast using the trained model on the test set (cross-basis matrix and hot/cold wave data)
  model_forecast <- B_DLNM_LC.forecast(train_model, 
                                       set$test$Roma_cb_test, 
                                       set$test$Roma_wave_test, 
                                       kt_model_type = "SARIMA", 
                                       forecast_step = forecast_step)
  
  # Return both the train model and the forecast for this iteration
  list(train_model = train_model, model_forecast = model_forecast,
       Roma_test = set$test$Roma_test)
})
```



# Forecasting mae function (calculate mae for each set by B-DLNM-LC vs baseline LC, and plot mae if necessary)
```{r}
expanding_window_forecast_mae <- function(set, plot_mae) {
  # Test values
  test_value <- exp(set$Roma_test)
  # B-DLNM-LC model forecast
  forecast_BDLNMLC <- exp(set$model_forecast$B_DLNM_LC_final.forecast$model.forecast)
  B_DLNM_LC_mae <- c(100*mae(test_value[,1], forecast_BDLNMLC[,1]),
                     100*mae(test_value[,2], forecast_BDLNMLC[,2]),
                     100*mae(test_value[,3], forecast_BDLNMLC[,3]),
                     100*mae(test_value[,4], forecast_BDLNMLC[,4]))
  
  # Baseline LC model forecast
  baseline_LC <- exp(set$model_forecast$baseline_LC.forecast)
  baseline_LC_mae <- c(100*mae(test_value[,1], baseline_LC[,1]),
                       100*mae(test_value[,2], baseline_LC[,2]),
                       100*mae(test_value[,3], baseline_LC[,3]),
                       100*mae(test_value[,4], baseline_LC[,4]))
  
  # Converged LC model forecast
  converged_LC <- exp(set$model_forecast$B_DLNM_LC_final.forecast$converged_LC.forecast)
  converged_LC_mae <- c(100*mae(test_value[,1], converged_LC[,1]),
                        100*mae(test_value[,2], converged_LC[,2]),
                        100*mae(test_value[,3], converged_LC[,3]),
                        100*mae(test_value[,4], converged_LC[,4]))
  # Return mae values and the plot
  return(list(B_DLNM_LC_mae = B_DLNM_LC_mae, baseline_LC_mae = baseline_LC_mae, converged_LC_mae = converged_LC_mae))
}

# Apply the function across all sets
forecast_mae <- lapply(expanding_window_result, function(set){
  expanding_window_forecast_mae(set, plot_mae = T) 
})
```

## Plot point forecast
```{r}
# Function to generate B-DLNM-LC forecasting plots
expanding_window_forecast_mae <- function(set, plot_mae, validation_index) {
  # Test values (keeping "Roma" for data extraction)
  test_value <- set$Roma_test
  
  # B-DLNM-LC model forecast
  forecast_BDLNMLC <- set$model_forecast$B_DLNM_LC_final.forecast$model.forecast
  
  # Baseline LC model forecast
  baseline_LC <- set$model_forecast$baseline_LC.forecast

  # Converged LC model forecast
  converged_LC <- set$model_forecast$B_DLNM_LC_final.forecast$converged_LC.forecast

  if (plot_mae) {
    # Prepare data for plotting
    age_groups <- c("20-64", "65-74", "75-84", "85+")
    plot_data <- data.frame(
      Time = rep(1:nrow(test_value), times = 4),
      Actual = as.vector(as.matrix(test_value)),
      B_DLNM_LC = as.vector(as.matrix(forecast_BDLNMLC)),
      Baseline_LC = as.vector(as.matrix(baseline_LC)),
      Age_Group = rep(age_groups, each = nrow(test_value))
    )
    
    # Reshape data to long format for ggplot
    plot_data_long <- plot_data %>%
      pivot_longer(cols = c("Actual", "B_DLNM_LC", "Baseline_LC"),
                   names_to = "Model", values_to = "Value")
    
    # Replace "Roma" with "Rome" in display titles
    display_region <- "Rome"

    # Create the plot with ggplot2
    p <- ggplot(plot_data_long, aes(x = Time, y = Value, color = Model)) +
      geom_line(size = 0.4) +  
      facet_wrap(~ Age_Group, scales = "free_y", ncol = 2) +
      scale_color_manual(values = c("black", "#D94738", "#1C61AB"),
                         labels = c("Actual Values", "B-DLNM-LC", "Baseline LC")) +
      labs(title = paste("Validation Set", validation_index),  
           x = "Week", y = "Mortality rate") +
      scale_y_continuous(n.breaks = 6) +  
      theme_minimal() +
      theme(
        panel.border = element_rect(color = "black", fill = NA, size = 0.5),  
        panel.grid = element_blank(),  
        strip.text = element_text(face = "bold", size = 9),  
        plot.title = element_text(hjust = 0.5, size = 12),  
        legend.position = "right",  
        legend.title = element_blank(),  
        plot.margin = margin(5, 5, 5, 5)  # Reduce white space
      )
    
    # Customize titles for each facet
    p <- p + facet_wrap(~ Age_Group, scales = "free_y", labeller = as_labeller(function(x) {
      paste("Age group", x)
    }))
    
    return(p)
  }
}

# Generate and collect plots
plots <- lapply(seq_along(expanding_window_result), function(i) {
  expanding_window_forecast_mae(expanding_window_result[[i]], plot_mae = TRUE, validation_index = i)
})

# Remove NULL values if some validation sets were missing data
plots <- Filter(Negate(is.null), plots)

# Arrange all 10 plots in a 2x5 grid
output_plot <- grid.arrange(
  grobs = plots, 
  ncol = 2,  # 2 columns
  nrow = 5,  # 5 rows
  top = textGrob("B-DLNM-LC Forecasting Results for Rome", 
                  gp = gpar(fontsize = 14, fontface = "bold"))
)

# Save the plot with updated filename (Rome instead of Roma)
ggsave(paste0("/Users/jiachengabelmin/Desktop/LC_DLNM/Figures/Forecasts/B_DLNM_LC/Rome_Forecast_5_8_104.pdf"),output_plot, width = 12, height = 8, dpi = 1000)

```
## Compute average Forecasting mse
```{r}
B_DLNM_LC_mae_values <- do.call(rbind, lapply(forecast_mae, function(set) set$B_DLNM_LC_mae))
baseline_LC_mae_values <- do.call(rbind, lapply(forecast_mae, function(set) set$baseline_LC_mae))
converged_LC_mae_values <- do.call(rbind, lapply(forecast_mae, function(set) set$converged_LC_mae))
age_groups <- c("20-64", "65-74", "75-84", "85+")
round((baseline_LC_ave_mae = colMeans(baseline_LC_mae_values)),4)
round((B_DLNM_LC_ave_mae = colMeans(B_DLNM_LC_mae_values)),4)
```

# Model residual analysis (consider the whole dataset)
## run the model
```{r}
# Smoothing cross-basis matrix on the whole dataset
varknots <- equalknots(UTCI_ext.Roma, fun = "ns", df = 3, degree = 3)
lagknots <- logknots(21, 3)
Roma_cb <- crossbasis(UTCI_ext.Roma, lag = 21, 
                        argvar = list(fun = "ns", knots = varknots), 
                        arglag = list(knots = lagknots, df = 3))

Roma.DLNMLC <- B_DLNM_LC(Roma, Roma_cb, Roma_wave, 
                           tol = 1e-2, max_iter = 20, region = "Roma")

# Model residual
Roma.residual = Roma - exp(Roma.DLNMLC$log_fitted)
```

## Normality test (Shapiro–Wilk test, Anderson–Darling test, and Jarque–Bera test)
In Shapiro–Wilk test, Anderson–Darling test, and Jarque–Bera test, H0: residual is Gaussian vs H1: residual is non-Gaussian. 
```{r}
# Shapiro–Wilk test
shapiro.p.value <- sapply(Roma.residual, function(x) shapiro.test(x[is.finite(x)])$p.value)
ifelse(shapiro.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Anderson–Darling test
ad.p.value <- sapply(Roma.residual, function(x) ad.test(x[is.finite(x)])$p.value)
ifelse(ad.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Jarque–Bera test
jarque.bera.p.value <- sapply(Roma.residual, function(x) jarque.bera.test(x[is.finite(x)])$p.value)
ifelse(jarque.bera.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
```
## Seasonality test on model residual
## Model residual - Roma
```{r}
# model residual
# Roma 
## QS test
Roma.residual.QS.p <- sapply(Roma.residual, function(residual) {
  round(qs(residual, freq = 52)$Pval, 3)
})

## Friedman test
Roma.residual.Friedman.p <- sapply(Roma.residual, function(residual) {
  round(fried(residual, freq = 52)$Pval, 3)
})

## Kruskall-Wallis test
Roma.residual.KW.p <- sapply(Roma.residual, function(residual) {
  round(kw(residual, freq = 52)$Pval, 3)
})
```
