---
title: "Section C. Figures of specific lagged effects via DLNM"
output: pdf_document
---

# Load packages
```{r section-c-packages, message=FALSE, warning=FALSE, results='hide'}
packages <- c(
  "readxl", "forecast", "dlnm", "splines", "Metrics", "demography",
  "tseries", "seastests"
)
invisible(lapply(packages, library, character.only = TRUE))
```

## Project root
```{r section-c-project-root, message=FALSE, warning=FALSE, results='hide'}
find_project_root <- function(start = getwd()) {
  path <- normalizePath(start, mustWork = FALSE)
  for (i in seq_len(8)) {
    if (dir.exists(file.path(path, "Code")) && dir.exists(file.path(path, "Data"))) {
      return(path)
    }
    parent <- dirname(path)
    if (identical(parent, path)) break
    path <- parent
  }
  stop("Set LC_DLNM_BASE_DIR to the project root.")
}

base_dir <- Sys.getenv("LC_DLNM_BASE_DIR")
if (!nzchar(base_dir)) base_dir <- find_project_root()

data_dir <- file.path(base_dir, "Data/Combined_data")
function_dir <- file.path(base_dir, "Code/Function")
```

## Load data
```{r section-c-load-data, message=FALSE, warning=FALSE, results='hide'}
Y20_64 <- read_xlsx(file.path(data_dir, "Y20_64_combined.xlsx"))
Y65_74 <- read_xlsx(file.path(data_dir, "Y65_74_combined.xlsx"))
Y75_84 <- read_xlsx(file.path(data_dir, "Y75_84_combined.xlsx"))
Y_GE85 <- read_xlsx(file.path(data_dir, "Y_GE85_combined.xlsx"))

age_columns <- c("Y20_64", "Y65_74", "Y75_84", "Y_GE85")
age_titles <- c("Age 20-64", "Age 65-74", "Age 75-84", "Age 85+")

region_specs <- list(
  Athens = list(death_col = 3, exposure_col = 9, source = "Attiki"),
  Lisbon = list(death_col = 4, exposure_col = 10, source = "Lisbon"),
  Rome = list(death_col = 6, exposure_col = 12, source = "Roma")
)
```

## Mortality and heat/cold wave inputs
```{r section-c-inputs, message=FALSE, warning=FALSE, results='hide'}
build_rate_matrix <- function(spec) {
  out <- cbind(
    Y20_64[[spec$death_col]] / Y20_64[[spec$exposure_col]],
    Y65_74[[spec$death_col]] / Y65_74[[spec$exposure_col]],
    Y75_84[[spec$death_col]] / Y75_84[[spec$exposure_col]],
    Y_GE85[[spec$death_col]] / Y_GE85[[spec$exposure_col]]
  )
  colnames(out) <- age_columns
  out * 52
}

build_wave_data <- function(region, spec) {
  out <- data.frame(
    hot_wave = Y20_64[[paste0(spec$source, "_hot_wave3")]],
    cold_wave = Y20_64[[paste0(spec$source, "_cold_wave3")]]
  )
  colnames(out) <- c(paste0(region, "_hot_wave3"),
                     paste0(region, "_cold_wave3"))
  out
}

dat_list <- lapply(region_specs, build_rate_matrix)
wave_list <- Map(build_wave_data, names(region_specs), region_specs)

Athens <- dat_list$Athens
Lisbon <- dat_list$Lisbon
Rome <- dat_list$Rome

Athens_wave <- wave_list$Athens
Lisbon_wave <- wave_list$Lisbon
Rome_wave <- wave_list$Rome
```

## Cross-basis matrices
```{r section-c-crossbasis, message=FALSE, warning=FALSE, results='hide'}
source(file.path(function_dir, "crossbasis_mixed_frequency.R"))

utci_list <- list(
  Athens = load_crossbasis_history(base_dir, "Athens", "Attiki"),
  Lisbon = load_crossbasis_history(base_dir, "Lisbon", "Lisbon"),
  Rome = load_crossbasis_history(base_dir, "Rome", "Roma")
)

make_crossbasis <- function(utci) {
  varknots <- equalknots(utci, fun = "ns", df = 3, degree = 3)
  lagknots <- logknots(21, fun = "ns", df = 3)
  crossbasis(
    utci,
    lag = 21,
    argvar = list(fun = "ns", knots = varknots),
    arglag = list(knots = lagknots, df = 3)
  )
}

cb_list <- lapply(utci_list, make_crossbasis)

Athens_cb <- cb_list$Athens
Lisbon_cb <- cb_list$Lisbon
Rome_cb <- cb_list$Rome
```

## Load model functions
```{r section-c-functions, message=FALSE, warning=FALSE, results='hide'}
files <- list.files(function_dir, pattern = "\\.R$", full.names = TRUE)
invisible(sapply(files, source))
```

## Fit DLNM--LC and DLNM--LL
```{r section-c-fit-models, message=FALSE, warning=FALSE, results='hide'}
invisible(capture.output({
  DLNM_LC_Athens_fit <- DLNM_LC(
    Athens, Athens_cb, wave_data = Athens_wave,
    tol = 1e-2, max_iter = 20, region = "Athens"
  )
  DLNM_LC_Lisbon_fit <- DLNM_LC(
    Lisbon, Lisbon_cb, wave_data = Lisbon_wave,
    tol = 1e-2, max_iter = 20, region = "Lisbon"
  )
  DLNM_LC_Rome_fit <- DLNM_LC(
    Rome, Rome_cb, wave_data = Rome_wave,
    tol = 1e-2, max_iter = 20, region = "Rome"
  )

  DLNM_LL_fit <- DLNM_LL(
    dat_list = dat_list,
    wave_list = wave_list,
    cb_list = cb_list,
    tol = 1e-2,
    max_iter = 20
  )
}))
```

## DLNM residual components
```{r section-c-dlnm-components, message=FALSE, warning=FALSE, results='hide'}
lc_temp <- list(
  Athens = as.data.frame(log(as.matrix(Athens)) - DLNM_LC_Athens_fit$final_LC$log_fitted),
  Lisbon = as.data.frame(log(as.matrix(Lisbon)) - DLNM_LC_Lisbon_fit$final_LC$log_fitted),
  Rome = as.data.frame(log(as.matrix(Rome)) - DLNM_LC_Rome_fit$final_LC$log_fitted)
)

ll_temp <- list(
  Athens = as.data.frame(log(as.matrix(Athens)) - DLNM_LL_fit$final_LL$log_fitted$Athens),
  Lisbon = as.data.frame(log(as.matrix(Lisbon)) - DLNM_LL_fit$final_LL$log_fitted$Lisbon),
  Rome = as.data.frame(log(as.matrix(Rome)) - DLNM_LL_fit$final_LL$log_fitted$Rome)
)

for (region in names(region_specs)) {
  colnames(lc_temp[[region]]) <- age_columns
  colnames(ll_temp[[region]]) <- age_columns
}
```

## Fit DLNM components for lagged-effect plots
```{r section-c-fit-dlnm-components, message=FALSE, warning=FALSE, results='hide'}
lc_centres <- list(
  Athens = c(Y20_64 = 22, Y65_74 = 22, Y75_84 = 20, Y_GE85 = 20),
  Lisbon = c(Y20_64 = 22, Y65_74 = 22, Y75_84 = 20, Y_GE85 = 20),
  Rome = c(Y20_64 = 18, Y65_74 = 20, Y75_84 = 22, Y_GE85 = 22)
)

ll_centres <- list(
  Athens = c(Y20_64 = 22, Y65_74 = 22, Y75_84 = 20, Y_GE85 = 20),
  Lisbon = c(Y20_64 = 22, Y65_74 = 22, Y75_84 = 20, Y_GE85 = 20),
  Rome = c(Y20_64 = 18, Y65_74 = 20, Y75_84 = 20, Y_GE85 = 20)
)

fit_dlnm_predictions <- function(temp_data, cb, wave_data, centres, region) {
  hot_col <- paste0(region, "_hot_wave3")
  cold_col <- paste0(region, "_cold_wave3")

  setNames(lapply(age_columns, function(age_col) {
    fit <- glm(
      temp_data[[age_col]] ~ cb + wave_data[[hot_col]] + wave_data[[cold_col]],
      family = gaussian()
    )
    crosspred(
      cb,
      fit,
      cen = centres[[age_col]],
      at = -5:40,
      model.link = "log"
    )
  }), age_columns)
}

lc_pred <- list(
  Athens = fit_dlnm_predictions(lc_temp$Athens, Athens_cb, Athens_wave,
                                lc_centres$Athens, "Athens"),
  Lisbon = fit_dlnm_predictions(lc_temp$Lisbon, Lisbon_cb, Lisbon_wave,
                                lc_centres$Lisbon, "Lisbon"),
  Rome = fit_dlnm_predictions(lc_temp$Rome, Rome_cb, Rome_wave,
                              lc_centres$Rome, "Rome")
)

ll_pred <- list(
  Athens = fit_dlnm_predictions(ll_temp$Athens, Athens_cb, Athens_wave,
                                ll_centres$Athens, "Athens"),
  Lisbon = fit_dlnm_predictions(ll_temp$Lisbon, Lisbon_cb, Lisbon_wave,
                                ll_centres$Lisbon, "Lisbon"),
  Rome = fit_dlnm_predictions(ll_temp$Rome, Rome_cb, Rome_wave,
                              ll_centres$Rome, "Rome")
)
```

## Plot helper
```{r section-c-plot-helper, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects <- function(predictions, ylim) {
  par(
    mfcol = c(2, 4),
    mar = c(5, 5, 2, 1),
    oma = c(1, 1, 3, 1),
    ps = 16,
    cex.lab = 1.2,
    cex.axis = 1.2,
    cex.main = 1.4,
    font.main = 2
  )

  for (age_col in age_columns) {
    plot(
      predictions[[age_col]], "slices",
      var = -5,
      col = 1,
      ylab = "RR",
      ci.arg = list(density = 50, lwd = 2, col = "#AFEEEE"),
      main = paste("UTCI = -5,", age_titles[[which(age_columns == age_col)]]),
      ylim = ylim
    )

    plot(
      predictions[[age_col]], "slices",
      var = 35,
      col = 1,
      ylab = "RR",
      ci.arg = list(density = 50, lwd = 2, col = "#AFEEEE"),
      main = paste("UTCI = 35,", age_titles[[which(age_columns == age_col)]]),
      ylim = ylim
    )
  }
}
```

## Figure C.1. DLNM--LC lagged effects for Athens
```{r section-c-lc-athens, fig.width=16, fig.height=8, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects(lc_pred$Athens, ylim = c(0.98, 1.05))
```

## Figure C.2. DLNM--LL lagged effects for Athens
```{r section-c-ll-athens, fig.width=16, fig.height=8, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects(ll_pred$Athens, ylim = c(0.98, 1.05))
```

## Figure C.3. DLNM--LC lagged effects for Lisbon
```{r section-c-lc-lisbon, fig.width=16, fig.height=8, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects(lc_pred$Lisbon, ylim = c(0.95, 1.2))
```

## Figure C.4. DLNM--LL lagged effects for Lisbon
```{r section-c-ll-lisbon, fig.width=16, fig.height=8, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects(ll_pred$Lisbon, ylim = c(0.95, 1.2))
```

## Figure C.5. DLNM--LC lagged effects for Rome
```{r section-c-lc-rome, fig.width=16, fig.height=8, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects(lc_pred$Rome, ylim = c(0.95, 1.1))
```

## Figure C.6. DLNM--LL lagged effects for Rome
```{r section-c-ll-rome, fig.width=16, fig.height=8, message=FALSE, warning=FALSE, results='hide'}
plot_lag_effects(ll_pred$Rome, ylim = c(0.95, 1.1))
```
