---
title: "HBY model"
author: "Jiacheng Min"
date: "`r Sys.Date()`"
output: pdf_document
---

## Load packages
```{r, warning = FALSE, output = FALSE}
packages <- c("readxl", "ggplot2", "forecast", "dplyr", "gridExtra", "writexl", 
              "tseries", "hwwntest", "dlnm", "splines", "ISOweek", "mgcv", 
              "Metrics", "demography", "seastests", "MTS", "reshape", "grid", 
              "tidyverse", "tidyr", "nortest")
invisible(lapply(packages, library, character.only = TRUE))
```


## Load data
```{r}
region_name = c("Attiki", "Lisbon", "Roma")
age = c('Y20_64', 'Y65_74', 'Y75_84', 'Y_GE85')
setwd("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/")
Y20_64 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y20_64_combined.xlsx")
Y65_74 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y65_74_combined.xlsx")
Y75_84 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y75_84_combined.xlsx")
Y_GE85 = read_xlsx("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Combined_data/Y_GE85_combined.xlsx")
``` 
## LC matrix (mortality rate)
```{r}
col_name = c("Y20_64", "Y65_74", "Y75_84", "Y_GE85")

Attiki = cbind(Y20_64[,3]/Y20_64[,9],Y65_74[,3]/Y65_74[,9], 
               Y75_84[,3]/Y75_84[,9], Y_GE85[,3]/Y_GE85[,9])
colnames(Attiki) = col_name

Lisbon = cbind(Y20_64[,4]/Y20_64[,10],Y65_74[,4]/Y65_74[,10],
               Y75_84[,4]/Y75_84[,10], Y_GE85[,4]/Y_GE85[,10])
colnames(Lisbon) = col_name

Roma = cbind(Y20_64[,6]/Y20_64[,12],Y65_74[,6]/Y65_74[,12], 
             Y75_84[,6]/Y75_84[,12], Y_GE85[,6]/Y_GE85[,12])
colnames(Roma) = col_name
```

## Store heat/cold wave data
```{r}
# Attiki
Attiki_wave = data.frame(cbind(Y20_64$Attiki_hot_wave3, Y20_64$Attiki_cold_wave3))
colnames(Attiki_wave) = c("Attiki_hot_wave3", "Attiki_cold_wave3")
# Lisbon
Lisbon_wave = data.frame(cbind(Y20_64$Lisbon_hot_wave3, Y20_64$Lisbon_cold_wave3))
colnames(Lisbon_wave) = c("Lisbon_hot_wave3", "Lisbon_cold_wave3")
# Roma
Roma_wave = data.frame(cbind(Y20_64$Roma_hot_wave3, Y20_64$Roma_cold_wave3))
colnames(Roma_wave) = c("Roma_hot_wave3", "Roma_cold_wave3")
```

# DLNM 
## Crossbasis matrix
```{r}
# before cross-basis
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Attiki.RData")
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Lisbon.RData")
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Roma.RData")
```


## Convert to standard weekly mortality (*52)
```{r}
Attiki = Attiki * 52
Lisbon = Lisbon * 52
Roma = Roma * 52
dat_list <- list(Attiki = Attiki, Lisbon = Lisbon, Roma = Roma)
wave_list <- list(Attiki = Attiki_wave, Lisbon = Lisbon_wave, Roma = Roma_wave)
cb_list <- list(Attiki = UTCI_ext.Attiki, Lisbon = UTCI_ext.Lisbon, Roma = UTCI_ext.Roma) 
### cb_list: for forecasting, we use ORIGINAL data! the parameter can be chosen in the forecasting function!
```

# Load functions
```{r}
files <- list.files("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Function/", pattern = "\\.R$", full.names = TRUE)
invisible(sapply(files, source))
```

## Backfitting DLNM-LL model with expanding window cross validation
```{r}
expanding_window_sets <- lapply(0:9, function(i) {
   create_train_test_sets(dat_list, wave_list, cb_list, train_size = 102 + i * 8, test_size = 180 + i * 8, lag = 21, var.df = 3, lag.df = 3, degree = 3)
})
names(expanding_window_sets) = paste0("set_", 1:10)
```

## Fitting + Forecasting  (Cross Validation)
```{r}
# Iterate over expanding_window_sets and train the model, forecast, and extract test values
expanding_window_result <- lapply(expanding_window_sets, function(set) {
    # Extract the number of rows for the test set
    h_step <- nrow(set$test_set$dat_list[[1]])

    # Train model
    train_model <- B_DLNM_LL(dat_list = set$train_set$dat_list,
                             cb_list = set$train_set$cb_list,
                             wave_list = set$train_set$wave_list,
                             tol = 1e-2, max_iter = 20)

    # Forecast values using the dynamically determined number of periods
    model_forecast <- B_DLNM_LL.forecast(train_model,
                                         cb_matrix_test = set$test_set$cb_list,
                                         wave_data_test = set$test_set$wave_list,
                                         forecast_step = 78)

    # Log-transform test values
    test_values <- lapply(set$test_set$dat_list, function(region) {
        lapply(region, function(age_group) {
            log(age_group)
        })
    })

    # Assign names to test values to match the structure of train model outputs
    names(test_values) <- names(train_model$final_LL$log_fitted)

    # Return a list containing the train model, model forecast, and test values
    list(train_model = train_model,
         model_forecast = model_forecast,
         test_values = test_values)
})

```

# Forecasting mae function (calculate mae for each set by B-DLNM-LL vs baseline LL, and plot mae if necessary)
```{r}
# Forecasting mae for each region each group (Backfitting DLNM-LL)
BDLNMLL_mae_list = lapply(expanding_window_result, function(set){
  ## Backfitting DLNM-LL
  forecast_BDLNMLL = lapply(set$model_forecast$B_DLNM_LL_final.forecast$model.forecast, function(region){
    as.data.frame(lapply(region, exp))
  })
  ## Test values
  test.values = lapply(set$test_values, function(region){
    as.data.frame(lapply(region, exp))
  })
  
  BDLNMLL_mae <- lapply(names(forecast_BDLNMLL), function(region) {
    sapply(names(forecast_BDLNMLL[[region]]), function(age_group) {
      mae(test.values[[region]][[age_group]], forecast_BDLNMLL[[region]][[age_group]])
    })
  })
  names(BDLNMLL_mae) <- names(forecast_BDLNMLL)
  BDLNMLL_mae <- lapply(BDLNMLL_mae, function(x) setNames(x, names(forecast_BDLNMLL[[1]])))
  return(BDLNMLL_mae)
})

# Forecasting mae for each region each group (Baseline LL)
baseline_LL_mae_list = lapply(expanding_window_result, function(set){
  ## Baseline LL
  forecast_baseline_LL = lapply(set$model_forecast$baseline_LL.forecast, function(region){
    as.data.frame(lapply(region, exp))
  })
  ## Test values
  test.values = lapply(set$test_values, function(region){
    as.data.frame(lapply(region, exp))
  })
  
  baseline_LL_mae <- lapply(names(forecast_baseline_LL), function(region) {
    sapply(names(forecast_baseline_LL[[region]]), function(age_group) {
      mae(test.values[[region]][[age_group]], forecast_baseline_LL[[region]][[age_group]])
    })
  })
  names(baseline_LL_mae) <- names(forecast_baseline_LL)
  baseline_LL_mae <- lapply(baseline_LL_mae, function(x) setNames(x, names(forecast_baseline_LL[[1]])))
  return(baseline_LL_mae)
})

extract_mae_line = function(mae_list){
  # Extract name from the list
  first_set <- names(mae_list)[1]
  first_region <- names(mae_list[[first_set]])[1]
  regions <- names(mae_list[[first_set]])
  age_groups <- names(mae_list[[first_set]][[first_region]])
  
  # Reconstruct the list
  BDLNMLL_mae_line  <- lapply(regions, function(region) {
    lapply(age_groups, function(age_group) {
      sapply(1:10, function(set_index) { # 10: for original cv, 12: by month
        mae_list[[paste0("set_", set_index)]][[region]][[age_group]]
      })
    })
  })
  names(BDLNMLL_mae_line) <- regions
  BDLNMLL_mae_line <- lapply(BDLNMLL_mae_line, function(age_group_list) {
    setNames(age_group_list, age_groups)
  })
}

BDLNMLL_mae_line = extract_mae_line(BDLNMLL_mae_list)
baseline_LL_mae_line = extract_mae_line(baseline_LL_mae_list)
```


## Plot point forecast
```{r}
# Function to generate B-DLNM-LL forecasting plots
expanding_window_forecast_mae_BDLNMLL <- function(set, plot_mae, validation_index, region) {
  # Extract region-specific data
  test_value <- set$test_values[[region]]  
  forecast_BDLNMLL <- set$model_forecast$B_DLNM_LL_final.forecast$model.forecast[[region]]
  baseline_LL <- set$model_forecast$baseline_LL.forecast[[region]]

  # Ensure `test_value` is a proper list of age groups before using cbind
  test_value_df <- if (is.list(test_value)) do.call(cbind, test_value) else as.matrix(test_value)
  forecast_BDLNMLL_df <- if (is.list(forecast_BDLNMLL)) do.call(cbind, forecast_BDLNMLL) else as.matrix(forecast_BDLNMLL)
  baseline_LL_df <- if (is.list(baseline_LL)) do.call(cbind, baseline_LL) else as.matrix(baseline_LL)
  
  # scale
  test_value_df = exp(test_value_df)
  forecast_BDLNMLL_df = exp(forecast_BDLNMLL_df)
  baseline_LL_df = exp(baseline_LL_df)
  
  # Ensure all dataframes have the same number of rows
  min_rows <- min(nrow(test_value_df), nrow(forecast_BDLNMLL_df), nrow(baseline_LL_df))
  test_value_df <- test_value_df[1:min_rows, , drop = FALSE]
  forecast_BDLNMLL_df <- forecast_BDLNMLL_df[1:min_rows, , drop = FALSE]
  baseline_LL_df <- baseline_LL_df[1:min_rows, , drop = FALSE]

  # Prepare data for plotting
  age_groups <- c("20-64", "65-74", "75-84", "85+")
  plot_data <- data.frame(
    Time = rep(1:min_rows, times = 4),
    Actual = as.vector(as.matrix(test_value_df)),
    B_DLNM_LL = as.vector(as.matrix(forecast_BDLNMLL_df)),
    Baseline_LL = as.vector(as.matrix(baseline_LL_df)),
    Age_Group = rep(age_groups, each = min_rows)
  )

  # Reshape data to long format for ggplot
  plot_data_long <- plot_data %>%
    pivot_longer(cols = c("Actual", "B_DLNM_LL", "Baseline_LL"),
                 names_to = "Model", values_to = "Value")

  # Replace display name for consistency
  display_region <- ifelse(region == "Attiki", "Attica", ifelse(region == "Roma", "Rome", region))

  # Create the plot with ggplot2
  p <- ggplot(plot_data_long, aes(x = Time, y = Value, color = Model)) +
    geom_line(size = 0.4) +  
    facet_wrap(~ Age_Group, scales = "free_y", ncol = 2) +  # 2x2 grid structure
    scale_color_manual(values = c("black", "#D94738", "#1C61AB"),  # Matching colors
                       labels = c("Actual Values", "B-DLNM-LL", "Baseline LL")) +
    labs(title = paste("Validation Set", validation_index),  
         x = "Week", y = "Mortality rate") +
    scale_y_continuous(n.breaks = 6) +  
    theme_minimal() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 0.5),  
      panel.grid = element_blank(),  
      strip.text = element_text(face = "bold", size = 9),  
      plot.title = element_text(hjust = 0.5, size = 12),  
      legend.position = "right",  
      legend.title = element_blank(),  
      plot.margin = margin(5, 5, 5, 5)  # Reduce white space
    )

  # Customize titles for each facet
  p <- p + facet_wrap(~ Age_Group, scales = "free_y", labeller = as_labeller(function(x) {
    paste("Age group", x)
  }))

  return(p)  # Return the plot for stacking
}

# Define regions for extraction (Attiki, Lisbon, Roma)
regions <- c("Attiki", "Lisbon", "Roma")

# Loop over each region and generate/save plots
for (selected_region in regions) {
  
  # Generate and collect plots for all validation sets in the selected region
  plots <- lapply(seq_along(expanding_window_result), function(i) {
    expanding_window_forecast_mae_BDLNMLL(expanding_window_result[[i]], plot_mae = TRUE, validation_index = i, region = selected_region)
  })

  # Remove NULL values if some validation sets were missing data
  plots <- Filter(Negate(is.null), plots)

  # Replace display name for consistency
  display_region <- ifelse(selected_region == "Attiki", "Attica", ifelse(selected_region == "Roma", "Rome", selected_region))

  # Arrange all plots in a grid (same structure as B-DLNM-LC)
  output_plot <- grid.arrange(
    grobs = plots, 
    ncol = 2,  # Keep 2 columns for consistency
    nrow = ceiling(length(plots) / 2),  # Adjust dynamically
    top = textGrob(paste("B-DLNM-LL Forecasting Results for", display_region), 
                   gp = gpar(fontsize = 14, fontface = "bold"))
  )

  # Save the plot with adjusted width & height
  ggsave(paste0("/Users/jiachengabelmin/Desktop/LC_DLNM/Figures/Forecasts/B_DLNM_LL/BDLNM_LL_Forecast_", display_region, "5_8_104.pdf"), 
         output_plot, width = 12, height = 8,  dpi = 1000)

  message(paste("Saved:", display_region, "forecasting results"))
}
```

## Compute average MAE
```{r}
baseline_ave_mae = lapply(baseline_LL_mae_line, function(region){
  lapply(region, function(age_group){
    return(mean(age_group))
  })
})

B_DLNM_LL_ave_mae = lapply(BDLNMLL_mae_line, function(region){
  lapply(region, function(age_group){
    return(mean(age_group))
  })
})

# Attiki
round(100*unlist(baseline_ave_mae$Attiki),4)
round(100*unlist(B_DLNM_LL_ave_mae$Attiki),4)
# Lisbon
round(100*unlist(baseline_ave_mae$Lisbon),4)
round(100*unlist(B_DLNM_LL_ave_mae$Lisbon),4)
# Roma
round(100*unlist(baseline_ave_mae$Roma),4)
round(100*unlist(B_DLNM_LL_ave_mae$Roma),4)
```

# Model residual analysis (consider the whole dataset)
## Run the model
```{r}
# Cross-basis matrix (smoothing)
# Attiki
varknots <- equalknots(UTCI_ext.Attiki, fun="ns",df = 3, degree = 3)
lagknots <- logknots(21,3)
Attiki_cb <- crossbasis(UTCI_ext.Attiki, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

# Lisbon
varknots <- equalknots(UTCI_ext.Lisbon,fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, 3)
Lisbon_cb <- crossbasis(UTCI_ext.Lisbon, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

# Roma
varknots <- equalknots(UTCI_ext.Roma,fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, 3)
Roma_cb <- crossbasis(UTCI_ext.Roma, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

cb_list <- list(Attiki = Attiki_cb, Lisbon = Lisbon_cb, Roma = Roma_cb) 

# run the model
DLNMLL = B_DLNM_LL(dat_list = dat_list, cb_list = cb_list, wave_list = wave_list, 
                   tol = 1e-2, max_iter = 20)

# Model residual
Attiki.residual = Attiki - exp(DLNMLL$log_fitted$Attiki)
Lisbon.residual = Lisbon - exp(DLNMLL$log_fitted$Lisbon)
Roma.residual = Roma - exp(DLNMLL$log_fitted$Roma)
```

## Normality test (Shapiro–Wilk test, Anderson–Darling test, and Jarque–Bera test)
In Shapiro–Wilk test, Anderson–Darling test, and Jarque–Bera test, H0: residual is Gaussian vs H1: residual is non-Gaussian.  
## Attiki
```{r}
# Shapiro–Wilk test
shapiro.p.value <- sapply(Attiki.residual, function(x) shapiro.test(x[is.finite(x)])$p.value)
ifelse(shapiro.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Anderson–Darling test
ad.p.value <- sapply(Attiki.residual, function(x) ad.test(x[is.finite(x)])$p.value)
ifelse(ad.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Jarque–Bera test
jarque.bera.p.value <- sapply(Attiki.residual, function(x) jarque.bera.test(x[is.finite(x)])$p.value)
ifelse(jarque.bera.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
```
## Lisbon
```{r}
# Shapiro–Wilk test
shapiro.p.value <- sapply(Lisbon.residual, function(x) shapiro.test(x[is.finite(x)])$p.value)
ifelse(shapiro.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Anderson–Darling test
ad.p.value <- sapply(Lisbon.residual, function(x) ad.test(x[is.finite(x)])$p.value)
ifelse(ad.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Jarque–Bera test
jarque.bera.p.value <- sapply(Lisbon.residual, function(x) jarque.bera.test(x[is.finite(x)])$p.value)
ifelse(jarque.bera.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
```

## Roma
```{r}
# Shapiro–Wilk test
shapiro.p.value <- sapply(Roma.residual, function(x) shapiro.test(x[is.finite(x)])$p.value)
ifelse(shapiro.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Anderson–Darling test
ad.p.value <- sapply(Roma.residual, function(x) ad.test(x[is.finite(x)])$p.value)
ifelse(ad.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
# Jarque–Bera test
jarque.bera.p.value <- sapply(Roma.residual, function(x) jarque.bera.test(x[is.finite(x)])$p.value)
ifelse(jarque.bera.p.value < 0.05, "Residual is non-Gaussian", "Residual is Guassian")
```
## Seasonality test on model residual
## Model residual - Attiki
```{r}
# model residual
# Attiki 
## QS test
Attiki.residual.QS.p <- sapply(Attiki.residual[age], function(residual) {
  round(qs(residual, freq = 52)$Pval, 3)
})

## Friedman test
Attiki.residual.Friedman.p <- sapply(Attiki.residual[age], function(residual) {
  round(fried(residual, freq = 52)$Pval, 3)
})

## Kruskall-Wallis test
Attiki.residual.KW.p <- sapply(Attiki.residual[age], function(residual) {
  round(kw(residual, freq = 52)$Pval, 3)
})
```

## Model residual - Lisbon
```{r}
# model residual
# Lisbon 
## QS test
Lisbon.residual.QS.p <- sapply(Lisbon.residual[age], function(residual) {
  round(qs(residual, freq = 52)$Pval, 3)
})

## Friedman test
Lisbon.residual.Friedman.p <- sapply(Lisbon.residual[age], function(residual) {
  round(fried(residual, freq = 52)$Pval, 3)
})

## Kruskall-Wallis test
Lisbon.residual.KW.p <- sapply(Lisbon.residual[age], function(residual) {
  round(kw(residual, freq = 52)$Pval, 3)
})
```

## Model residual - Roma
```{r}
# model residual
# Roma 
## QS test
Roma.residual.QS.p <- sapply(Roma.residual[age], function(residual) {
  round(qs(residual, freq = 52)$Pval, 3)
})

## Friedman test
Roma.residual.Friedman.p <- sapply(Roma.residual[age], function(residual) {
  round(fried(residual, freq = 52)$Pval, 3)
})

## Kruskall-Wallis test
Roma.residual.KW.p <- sapply(Roma.residual[age], function(residual) {
  round(kw(residual, freq = 52)$Pval, 3)
})
```

