---
title: "Simulation_rcp85"
author: "Jiacheng Min"
date: "`r Sys.Date()`"
output: html_document
---

# Load packages
```{r, warning = FALSE, output = FALSE}
packages <- c("readxl", "ggplot2", "forecast", "dplyr", "gridExtra", "writexl", 
              "tseries", "hwwntest", "dlnm", "splines", "ISOweek", "mgcv", 
              "Metrics", "demography", "seastests", "MTS", "reshape2", "astsa",
              "MASS", "patchwork")
lapply(packages, library, character.only = TRUE)
```

# Load functions
```{r}
files <- list.files("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Function/", pattern = "\\.R$", full.names = TRUE)
sapply(files, source)
```

# Load historical UTCi and mortality data
```{r}
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Historical_data_aggregation/Historical_data.RData")
```

# Load historical crossbasis matrix data
```{r}
# *** Before 2D smoothing ***
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Attiki.RData")
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Lisbon.RData")
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Code/Crossbasis_matrix/cb_Roma.RData")

# Attiki
varknots <- equalknots(UTCI_ext.Attiki, fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, fun = "ns", df = 3)
Attiki_cb <- crossbasis(UTCI_ext.Attiki, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

# Lisbon
varknots <- equalknots(UTCI_ext.Lisbon,fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, fun = "ns", df = 3)
Lisbon_cb <- crossbasis(UTCI_ext.Lisbon, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

# Roma
varknots <- equalknots(UTCI_ext.Roma,fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, fun = "ns", df = 3)
Roma_cb <- crossbasis(UTCI_ext.Roma, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

cb_list <- list(Attiki = Attiki_cb, Lisbon = Lisbon_cb, Roma = Roma_cb)
rm(UTCI_ext.Attiki, UTCI_ext.Lisbon, UTCI_ext.Roma)
rm(Attiki_cb, Lisbon_cb, Roma_cb)
```

# Load future UTCI data and future corssbasis matrix data (rcp85)
```{r}
# *** Before 2D smoothing ***
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/Attiki_cb_rcp85.RData")
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/Lisbon_cb_rcp85.RData")
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/Roma_cb_rcp85.RData")


# Attiki
varknots <- equalknots(Attiki_prep_cb_rcp85, fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, fun = "ns", df = 3)
Attiki_cb.rcp85 <- crossbasis(Attiki_prep_cb_rcp85, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

# Lisbon
varknots <- equalknots(Lisbon_prep_cb_rcp85,fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, fun = "ns", df = 3)
Lisbon_cb.rcp85 <- crossbasis(Lisbon_prep_cb_rcp85, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))

# Roma
varknots <- equalknots(Roma_prep_cb_rcp85,fun="ns",df = 3, degree = 3)
lagknots <- logknots(21, fun = "ns", df = 3)
Roma_cb.rcp85 <- crossbasis(Roma_prep_cb_rcp85, lag=21, 
                                  argvar=list(fun="ns", knots = varknots), 
                                  arglag=list(knots=lagknots,df = 3))
cb_list_rcp85 = list(Attiki = Attiki_cb.rcp85, Lisbon = Lisbon_cb.rcp85, Roma = Roma_cb.rcp85)
rm(Attiki_prep_cb_rcp85, Lisbon_prep_cb_rcp85, Roma_prep_cb_rcp85)
rm(Attiki_cb.rcp85, Lisbon_cb.rcp85, Roma_cb.rcp85)
```

# Load Future Heat/Cold Wave Data (rcp85)
```{r}
load("/Users/jiachengabelmin/Desktop/LC_DLNM/Data/Simulation_data/UTCI/rcp85/wave_list_rcp85.RData")
```

# Fit Backfitting DLNM-LL model
```{r}
B_DLNM_LL_fitting = B_DLNM_LL(dat_list = dat_list, cb_list = cb_list, wave_list = wave_list, tol = 1e-2, max_iter = 20)
nsim = 10000 # the number of simulations
```


# Fit SARIMA model on K(t) and kappa(t,i)
```{r}
K.ts.model = kappa_fit(B_DLNM_LL_fitting, 104)
K_fit_list = list(common_fit = K.ts.model$final_kt$kt.fit$Common_fit,
                  Attiki_fit = K.ts.model$final_kt$kt.fit$Attiki_fit,
                  Lisbon_fit = K.ts.model$final_kt$kt.fit$Lisbon_fit, 
                  Roma_fit = K.ts.model$final_kt$kt.fit$Roma_fit)
```

# Simulation of K(t) and kappa(t,i) (Mannualy extract coefficeints and orders)
- Start at 2020/01/01. To match the time period 2031-2045, we simulate 26 years (1352 weeks) to cover the time period for 10000 paths.
```{r}
set.seed(111)
K_sim_list <- list(
  common_K.sim = data.frame(replicate(n = nsim, sarima.sim(n = 52*26, ar = 0.757553149, d = 0, ma = c(-0.319981635, 0.009969013), sar = 0, D = 0, sma = 0.106674692, S = 52))),
  Attiki_k.sim =  data.frame(replicate(n = nsim, sarima.sim(n = 52*26, ar = c(0.40358338, 0.25462489), d = 0, ma = c(-0.04433412, -0.30213904), sar = 0, D = 0, sma = 0.14377324, S = 52))),
  Lisbon_k.sim =  data.frame(replicate(n = nsim, arima.sim(n = 52*26, model = list(ar = 0.3361887)))),
  Roma_k.sim = data.frame(replicate(n = nsim, sarima.sim(n = 52*26, ar = c(1.53866985, -0.55864782), d = 0, ma = c(-1.38872694, 0.43411636), sar = -0.06684183, D = 0, sma = 0, S = 52)))
)
```

# Converged LL 
```{r}
# Simulated parameters
Axs = B_DLNM_LL_fitting$final_LL$Ax # list
Bxs = B_DLNM_LL_fitting$final_LL$Bx 
bxs = B_DLNM_LL_fitting$final_LL$bx # list 
kts = K_sim_list[-1]
Kts = K_sim_list[[1]]

## B(x) %*% K(t) simulation
Bxs_Kts_list = lapply(1:ncol(Kts), function(ind) {
  outer(Kts[, ind], Bxs, FUN = "*")
})
## b(x, i) %*% k(t) simulation
bxs_kts_list = list()
for (ind in 1:length(kts)){
  bxs_kts_list[[ind]] = lapply(1:ncol(kts[[ind]]), function(col) {
  outer(kts[[ind]][, col], bxs[[ind]], FUN = "*")
})
}
names(bxs_kts_list) = c("Attiki", "Lisbon", "Roma")
sim.BKbk <- lapply(bxs_kts_list, function(sublist) {
  Map(function(x, y) x + y, sublist, Bxs_Kts_list)
})

converged_LL.sim.rcp85 = Map(function(sim_sublist, axs_values) {
  # sublist
  lapply(sim_sublist, function(df) {
    # add A(x, i)
    sweep(df, 2, axs_values, FUN = "+")
  })
}, sim.BKbk, Axs)

converged_LL.sim.rcp85 <- lapply(converged_LL.sim.rcp85, function(region) {
  n_age_groups <- ncol(region[[1]])
  age_group_sims <- vector("list", n_age_groups)
  names(age_group_sims) <- c("Y20_64", "Y65_74", "Y75_84", "Y_GE85")
  for (sim_index in seq_along(region)) {
    for (age_group_index in seq_len(n_age_groups)) {
      if (is.null(age_group_sims[[age_group_index]])) {
        age_group_sims[[age_group_index]] <- list()
      }
      age_group_sims[[age_group_index]][[sim_index]] <- region[[sim_index]][, age_group_index]
    }
  }
  age_group_sims
})
```

# DLNM Coefficient Bootstrap Simulation
```{r}
dlnm_coef = B_DLNM_LL_fitting$dlnm_coef
dlnm_vcov = B_DLNM_LL_fitting$dlnm_vcov
boot_coef <- lapply(names(dlnm_coef), function(region) {
  lapply(names(dlnm_coef[[region]]), function(iteration) {
    lapply(names(dlnm_coef[[region]][[iteration]]), function(age_group) {
      coef_vec <- dlnm_coef[[region]][[iteration]][[age_group]]
      vcov_mat <- dlnm_vcov[[region]][[iteration]][[age_group]]
      mvrnorm(n = nsim, mu = coef_vec, Sigma = vcov_mat)
    })
  })
})
names(boot_coef) <- names(dlnm_coef)
for (region in names(boot_coef)) {
  names(boot_coef[[region]]) <- names(dlnm_coef[[region]])
  for (iteration in names(boot_coef[[region]])) {
    names(boot_coef[[region]][[iteration]]) <- names(dlnm_coef[[region]][[iteration]])
  }
}
```

# DLNM Part Simulation
## Crossbasis matrix simulation results
```{r}
## (1) dlnm_coef %*% smoothing crossbasis matrix (each iteration each age group)
cb_matrix.sim = lapply(seq_along(boot_coef), function(i) { 
  region_cb <- cb_list_rcp85[[i]]  # crossbasis matrix
  
  lapply(boot_coef[[i]], function(iteration) {
    lapply(iteration, function(age_group) {
      # 100 simulations
      lapply(seq_len(nrow(age_group)), function(bootstrap_index) {
        coef <- age_group[bootstrap_index, ]
        apply(region_cb, 1, function(row) {
          sum(row * coef[2:(length(coef) - 2)]) + coef[1]
        })
      })
    })
  })
})
names(cb_matrix.sim) = names(boot_coef)

## (1.1) Sum up all forecast values from DLNMs.
cb_matrix.sum.sim <- lapply(cb_matrix.sim, function(region) {
  age_group_sums <- list()
  for (iteration in region) {
    for (age_group_name in names(iteration)) {
      if (!age_group_name %in% names(age_group_sums)) {
        age_group_sums[[age_group_name]] <- iteration[[age_group_name]]
      } else {
        age_group_sums[[age_group_name]] <- mapply(
          `+`, 
          age_group_sums[[age_group_name]], 
          iteration[[age_group_name]], 
          SIMPLIFY = FALSE
        )
      }
    }
  }
  age_group_sums
})
```

## Heat/Cold Wave Simulation Results
```{r}
## (2) dlnm_coef %*% hot/cold wave term (each iteration each age group)
wave.sim = lapply(seq_along(boot_coef), function(i) { 
  region_wave <- wave_list_rcp85[[i]] # regional wave data
  lapply(boot_coef[[i]], function(iteration) {
    lapply(iteration, function(age_group) {
      # 100 simulations
      lapply(seq_len(nrow(age_group)), function(bootstrap_index) {
        coef <- age_group[bootstrap_index, ]
        apply(region_wave, 1, function(row) {
          sum(row * coef[(length(coef)-1):length(coef)])
        })
      })
    })
  })
})
names(wave.sim) = names(boot_coef)
## (2.1) Sum up all forecast values from DLNMs.

wave.sum.sim <- lapply(wave.sim, function(region) {
  age_group_sums <- list()
  for (iteration in region) {
    for (age_group_name in names(iteration)) {
      if (!age_group_name %in% names(age_group_sums)) {
        age_group_sums[[age_group_name]] <- iteration[[age_group_name]]
      } else {
        age_group_sums[[age_group_name]] <- mapply(
          `+`, 
          age_group_sums[[age_group_name]], 
          iteration[[age_group_name]], 
          SIMPLIFY = FALSE
        )
      }
    }
  }
  age_group_sums
})
```

## Summation of Crossbasis Matrix And Wave Results
```{r}
dlnm.sum.sim <- Map(function(cb_region, wave_region) {
  age_group_sums <- lapply(names(cb_region), function(age_group_name) {
    cb_age_group <- cb_region[[age_group_name]]
    wave_age_group <- wave_region[[age_group_name]]
    
    mapply(function(cb_row, wave_row) {
      cb_row + wave_row
    }, cb_age_group, wave_age_group, SIMPLIFY = FALSE)
  })
  
  names(age_group_sums) <- names(cb_region)
  age_group_sums
}, cb_matrix.sum.sim, wave.sum.sim)
```

## Summation of Converged LL Part And DLNM Part
```{r}
final_LL.sim.rcp85 <- Map(function(region_converged, region_dlnm) {
  Map(function(age_group_converged, age_group_dlnm) {
    mapply(function(sim_converged, sim_dlnm) {
      Lc <- length(sim_converged)   # original length
      Ld <- length(sim_dlnm)         # length of the DLM adjustment

      # un‐adjusted “head” of the series
      unchanged_part <- sim_converged[1:(Lc - Ld)]
      # the tail we want to adjust
      adjusted_part  <- sim_converged[(Lc - Ld + 1):Lc] + sim_dlnm

      # stitch back together, then force back to length Lc
      head(c(unchanged_part, adjusted_part), Lc)
    }, age_group_converged, age_group_dlnm, SIMPLIFY = FALSE)
  }, region_converged, region_dlnm)
}, converged_LL.sim.rcp85, dlnm.sum.sim)

age_group_names <- c("Y20_64", "Y65_74", "Y75_84", "Y_GE85")
final_LL.sim.rcp85 <- lapply(final_LL.sim.rcp85, function(region) {
  lapply(seq_along(region[[1]]), function(sim_index) {
    sim_data <- sapply(region, function(age_group) age_group[[sim_index]])
    data.frame(sim_data, check.names = FALSE) |> 
      setNames(age_group_names)
  })
})
```

# Add Error Terms, Store Converged LL And B-DLNM-LL Simulation Results 
```{r}
# Generate error terms
generate_error_terms <- function(original_data, sd = 0.0001) {
  # original_data: Reference structure for generating error terms
  # sd: Standard deviation of the error terms
  lapply(original_data, function(region) {
    lapply(region, function(simulation) {
      # Generate random error terms matching the dimensions of each simulation
      as.data.frame(matrix(rnorm(n = prod(dim(simulation)), mean = 0, sd = sd), 
                           nrow = nrow(simulation), ncol = ncol(simulation),
                           dimnames = list(NULL, colnames(simulation))))
    })
  })
}

# Generate error terms with the same structure as `final_LL.sim.rcp85` and `converged_LL.sim.rcp85`
error_terms <- generate_error_terms(final_LL.sim.rcp85, sd = 0.0001)

# Add error terms to `final_LL.sim.rcp85`
final_LL.sim.rcp85 <- Map(function(data, error) {
  lapply(seq_along(data), function(i) {
    Map(`+`, data[[i]], error[[i]])  # Add error terms to corresponding data
  })
}, final_LL.sim.rcp85, error_terms)

# Rename the simulations within each region
final_LL.sim.rcp85 <- lapply(final_LL.sim.rcp85, function(region) {
  names(region) <- paste0("simulation_", seq_along(region))
  region
})

# Data Structure
final_LL.sim.rcp85 <- lapply(final_LL.sim.rcp85, function(region) {
  lapply(region, function(simulations) {
    as.data.frame(do.call(cbind, simulations))
  })
})

# Add error terms to `converged_LL.sim.rcp85`
converged_LL.sim.rcp85 <- Map(function(data, error) {
  lapply(seq_along(data), function(i) {
    Map(`+`, data[[i]], error[[i]])  # Add error terms to corresponding data
  })
}, converged_LL.sim.rcp85, error_terms)

# Rename the simulations within each region and rename age groups
converged_LL.sim.rcp85 <- lapply(converged_LL.sim.rcp85, function(region) {
  # Rename the age groups inside each region
  age_group_names <- c("Y20_64", "Y65_74", "Y75_84", "Y_GE85")  # Ensure correct names for age groups
  names(region) <- age_group_names
  
  # Rename the simulations inside each age group
  lapply(region, function(age_group) {
    names(age_group) <- paste0("simulation_", seq_along(age_group))
    age_group
  })
})

converged_LL.sim.rcp85 <- lapply(converged_LL.sim.rcp85, function(region) {
  age_group_names <- names(region)
  num_simulations <- length(region[[1]])
  simulations <- lapply(seq_len(num_simulations), function(sim_index) {
    sim_data <- sapply(age_group_names, function(age_group) {
      region[[age_group]][[sim_index]] 
    })
    data.frame(sim_data) |> setNames(age_group_names)  
  })
  names(simulations) <- paste0("simulation_", seq_along(simulations))
  simulations
})

save(final_LL.sim.rcp85, file = "/Users/jiachengabelmin/Desktop/LC_DLNM/Results/Simulated_mortality_rates/final_LL_rcp85.RData")
```
